---
title: "Analyzing Load Test Logs"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "80%",
  fig.align = "center",
  fig.width = 8, fig.height = 6
)
```

# Log output

There are a few types of log files generated by a `shinycannon` recording session.

* `./recording.log`
  * Copy of the original recording used
  * This file is a file comprised of JSON lines that capture all of the HTTP(s) and WebSocket traffic that occured during the recording
* `./detail.log`
  * Log output to help debug errors
* `./sessions/*.csv`
  * Collection of files for each session number, simulated user id, and user session
  * One file is generated for each simulated session
  * The file name format displays `SESSION_USER_USERSESSION.csv`
  * Lines starting with `#` should be ignored when processing

These outputs are combined when creating tidy shinyloadtest data.

# Tidy Data

Shiny cannon recording sessions can be loaded into R using the `tidy_loadtest` function.


```{r, eval = FALSE}
df <- shinyloadtest::tidy_loadtest(
  `Run 1` = "./PATH/TO/RUN_1",
  `Run 2` = "./PATH/TO/RUN_2",
  `Run 3` = "./PATH/TO/RUN_3"
)
```

The `shinyloadtest::tidy_loadtest` function uses these files to generate a data frame with loadtest results. The resulting data frame contains the following columns:

* `run`
  * The name of the recording session. This would be `Run 1` in the example above
* `session_id`
  * For the given `run`, an incrementing integer value of the session being recorded. Starts at 0.
* `user_id`
  * For the given `run`, which simulated user is performing the work. Starts at 0.
* `iteration`
  * An incrementing integer value of the session iteration of a particular `user`. Starts at 0.
* `input_line_number`
  * The line number corresponding to the event in the `recording.log` file.
* `event`
  * The web event being performed.  Could be one of the following:
    * `REQ_HOME` - The initial HTTP request for the application to load the Homepage
    * `REQ_GET` - Request a supporting file (JavaScript / CSS)
    * `REQ_TOK` - Request a Shiny token
    * `REQ_SINF` - Request WebSocket information
    * `REQ_POST` - Perform a POST query, such as uploading part of a file
    * `WS_RECV_BEGIN_UPLOAD` - A file upload is being requested
    * `WS_OPEN` - Open a new WebSocket connection
    * `WS_RECV_INIT` - Initialize a new WebSocket
    * `WS_SEND` - Send information from the Shiny server to the browser
    * `WS_RECV` - Send information from the browser to the Shiny server
    * `WS_CLOSE` - Close the WebSocket connection
* `start`
  * Start time of the event relative to the beginning of the `run`'s maintenance period
* `end`
  * End time of the event relative to the beginning of the `run`'s maintenance period
* `time`
  * Total ellapsed time of the event
* `concurrency`
  * A number of events that are being processed concurrently
* `maintenance`
  * A boolean determining whether or not all simulated users are execution a session
* `label`
  * A human readible event name
* `json`
  * The parse JSON provided in the `recording.log` file.  If the field `message` exists, a `message_parsed` field is added containing a parsed form of the WebSocket's JSON message content.


> Normally, users will not need to worry about specific files, but should be able to use the tidy data frame for further analysis.

They key idea is that each row of the tidy results data frame represents an event. Events include the initial page load, user inputs and their associated outputs, and user wait times. You'll want to compare the duration of event times across different load tests.


# Website Anatomy

Websites are comprised of two main parts:

* Main [HTML](https://en.wikipedia.org/wiki/HTML) document
* Supporting files, such as images, [CSS](https://en.wikipedia.org/wiki/Cascading_Style_Sheets) files (webpage style), and [JavaScript](https://en.wikipedia.org/wiki/JavaScript) files (page interactivity).

The main HTML document must be processed to determine which supporting files to retrieve.  This is why you will always see the first event as `Requesting the Homepage`.  Many (~ 10) of the following events will be to request specific CSS or JavaScript files.  After loading the CSS files, the browser will then style the webpage.

Once the JavaScript files have finished loading, communication to a data server can begin.  For Shiny applications, this involves communicating over a [WebSocket](https://en.wikipedia.org/wiki/WebSocket).  A WebSocket is a single, constant line of communication between the Shiny server and the Shiny application in the browser.  The WebSocket will notify the Shiny server about which input values have changed by a user and will also notify the Shiny application about which outputs need to been updated.  This communication process will continue until the WebSocket is closed, typically the last line of the `recording.log` file.

As an example, the basic [`01_hello` gallery shiny app](https://gallery.shinyapps.io/001-hello/) loads the homepage, loads more than 20 supporting CSS and JavaScript files, and then begins WebSocket communication with the supporting Shiny server.


# Report Output

The data provided by `shinyloadtest::tidy_loadtest` can be used to directly create a stand alone HTML report.  This report contains six major sections of diagnostic information.

```{r, include = FALSE}
df <- shinyloadtest::tidy_loadtest(
  `1 user` = "../test_sessions/demo1",
  `4 users` = "../test_sessions/demo4",
  `16 users` = "../test_sessions/demo16"
)
# ignore <- shinyloadtest::make_report(df, "report.html", verbose = FALSE, open_browser = FALSE)
```
```{r, eval = FALSE}
df <- shinyloadtest::tidy_loadtest(
  "1 user" = "./demo1",
  "4 users" = "./demo4",
  "16 users" = "./demo16"
)
shinyloadtest::make_report(df, "report.html")
```

## Sessions

The `Sessions` tab of the shinyloadtest report displays the each simulated session performed by each user.  Each run displays their respective [gantt chart](https://en.wikipedia.org/wiki/Gantt_chart) by clicking on the named buttons near the top of the webpage.

```{r, echo = FALSE, fig.height = 3}
suppressPackageStartupMessages(library(dplyr, quiet = TRUE))
suppressPackageStartupMessages(library(shinyloadtest, quiet = TRUE))
df %>%
  filter(run == "4 users") %>%
  plot_gantt()
```

Sessions that are desaturated in color are sessions that have started before (warmup) or ended after (cooldown) the maintenance period.  The maintenance period is marked with vertical dotted lines spanning every simulated user.

In the example above, four users were stagger started and simulated for 10 minutes. The majority of the time is spent calculating new values for the outputs in the recording.  The session gantt chart below displays the same recording but with sixteen users.  

```{r, echo = FALSE, fig.height = 5}
df %>%
  filter(run == "16 users") %>%
  plot_gantt()
```

Both runs were executed on a single processor.  With very high resource contention, a larger proportion of time is spent calculating in the `16 users` run.

As more users are placed in an application, the average execution time might go up.  The `4 users` case had 'Simulated User # 0' start almost immediately after mantenance mode began.  However, the `16 users` run should have had a larger stagger start than what was performed.  Most of the other users did not start a new session until 40-50 seconds after 'Simulated User #16' began.  This is not uniform server load behavior.  

It is recommended to uniformly stagger start the user sessions. This can be guestimated by:
1. Record a run
2. Get an average session duration time
3. Divide that session duration by the number of simulated users
4. Set the new value as the stagger start value for a new `shinycannon` recording

```{r}
library(dplyr)
df %>% 
  group_by(run, session_id, user_id) %>% 
  summarise(duration = max(end) - min(start)) %>%
  group_by(run) %>% 
  summarize(
    avg_duration = mean(duration), 
    stagger_start = avg_duration / length(unique(user_id)))
```


## Session Duration

The `Session Duration` tab displays all maintenance period sessions ordered by total duration vertically from fastest to slowest.  

```{r, echo = FALSE}
df %>% filter(run == "4 users") %>% plot_gantt_duration(cutoff = attr(df, "recording_duration"))
```

Ideally, most of the sessions should be completed in about the same amount of time as the recording.log execution time.  The vertical red line line defaults to the `recording.log` duration, but can be set when calling `make_report(df, duration_cutoff = 10)`.  In the example above, most sessions finish around the same amount of time as the `recording.log` duration.

It is good to have to have each session duration (row) look like the next.  If certain rows are much wider than the others, server resource contention might be good first suspect for the delay.  Looking at the `16 user` run, the same recording is on average 4x slower than the `recording.log` duration with one session taking twice as long as the average `16 user` session.  

```{r, echo = FALSE}
df %>% filter(run == "16 users") %>% plot_gantt_duration(cutoff = attr(df, "recording_duration"))
```


## Event Waterfall

The `Event Waterfall` tab displays an event waterfall of each recording session.  In an ideal world with no calculation time, all lines would be vertical.  In a practical world, all of the lines should be parallel to each other.  This would mean that the calculation times are consistent and are not behaving erradically.

Looking at the 4 user run below, the lines are fairly uniform in shape and are mostly parallel to eachother.

```{r, echo = FALSE, fig.height=5}
df %>% filter(run == "4 users") %>% plot_timeline()
```

If we look at another run where a file was uploaded and many POST requests were made for a DT table, there are many horizontal lines that are not parallel to each other. This happens when execution for that step took **much** more time than the other sessions.  The lines that jut horizontally indicate erratic behavior.

```{r, echo = FALSE, fig.height=10}
upload <- tidy_loadtest(`upload-dt` = "../test_sessions/upload-dt/", verbose = FALSE)
upload %>% plot_timeline()
```


## Latency

The `Latency` tab displays two plots.  The first plot displays the total HTTP time for each session.  The second plot displays the maximum calculation time.  Both plots display a cutoff value that can be set when calling `make_report(df, http_latency_cutoff = 5, max_websocket_cutoff = 20)`. Similar to the `Sessions` tab, sessions that are desaturated in color are sessions that started before (warmup) or ended after (cooldown) the maintenance period.

The total HTTP latency plot helps to diagnose how long a user has to wait before a website has appeared on their browser.  Ideally, this should be no time at all, as the server is just providing static files.

```{r, echo = FALSE, fig.height=7}
df %>% plot_http_latency()
```

The maximum WebSocket latency plot helps to diagnose the maximum amount of time a user has to wait before an input is updated or output is recalculated.  Ideally, this should be shorter than our attention spans.

```{r, echo = FALSE, fig.height=7}
df %>% plot_websocket_latency()
```

In the example above, the `16 users` run has a very strong pattern in it's maximum calculation time.  Knowing that `16 users` run has the same recording as the `1 user` and `4 users` runs, it would appear that the `16 users` server is under too much load as the same calculations could be computed within 1 second (as in the `1 user` run).  


## Event Duration

The `Event Duration` tab has three to four tabs available to choose from:

* Slowest max time - Event plots are arranged by the slowest maximum time within each plot
* Slowest min time - Event plots are arranged by the slowest minimum time within each plot
* Largest mean difference - Event plots are arranged by the largest difference in the run's mean event time. This option is available if more than one run is being displayed.
* Data table - A tabular display of the slowest max time, slowest min time, and (possibly) largest mean difference

```{r, echo = FALSE}
df %>% filter(input_line_number == 1) %>% plot_time_boxplot()
```

Ideally, the boxplots should not differ for any event and are close to 0 seconds ellapsed.  As resource contention goes up, event time might go up.  Hopefully, there are no extreme outliers above the center mass of the boxplot as shown below.

```{r, echo = FALSE}
df %>% filter(input_line_number == 4) %>% plot_time_boxplot()
```


## Event Concurrency


The `Event Concurrency` tab has four tabs available to choose from:

* Largest slope - Event plots are arranged by the largest slope magnitude found when fitting a linear model to each run
* Largest intercept - Event plots are arranged by the largest intercept magnitude found when fitting a linear model to each run
* Largest error - Event plots are arranged by the largest error magnitude found when fitting a linear model to each run
* Data table - A tabular display of the raw slope, raw intercept, and raw error values

Ideally, we want a flat linear model with no slope.  This means that the calculations of our shiny application will scale as user concurrency goes up.  

```{r, echo = FALSE}
df %>% filter(input_line_number == 16) %>% plot_concurrency_time()
```


However, if a strong linear trend exists as concurrency is increased, users will be waiting longer as more users are accessing the shiny application.

```{r, echo = FALSE}
df %>% filter(input_line_number == 34) %>% plot_concurrency_time()
```


<!-- # Tips -->

<!-- * It is always good to establish a baseline or an __approved__ run when comparing against potential server configuration runs.  This will give you a baseline to compare against.   -->

<!-- * Try to only alter one server configuration component at a time when deviating from your baseline run.  This helps pinpoint exactly what caused a speedup or slowdown in the application response time.   -->
