---
title: "Installing and running shinycannon"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing and running shinycannon}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shinyloadtest)
```

## Installation

`shinycannon` requires custom installation that differs from platform to platform.

### Linux

Depending on your distribution, `shinycannon` can be installed using one of the following packages.

```{r, include=FALSE}
library(readr)
library(magrittr)
library(dplyr)
library(knitr)

# Release urls are updated in Rituals GHA
shinycannon_files <-
  read_csv(
    "RELEASE_URLS.csv",
    comment = "#",
    col_types = list(.default = col_character())
  ) %>%
  mutate(`Download Link` = sprintf("<a href='%s'>%s</a>", url, file))

package_types <- tribble(
  ~platform,  ~Distribution,          ~command_str,
  "deb",      "Ubuntu/Debian",        "<code>sudo dpkg -i %s</code>",
  "rpm_rh",   "Redhat/Fedora/CentOS", "<code>sudo yum install -y %s</code>",
  "rpm_suse", "SUSE",                 "<code>zypper ----no-gpg-checks install -y %s</code>"
)

linux_packages <- shinycannon_files %>%
  right_join(package_types, by = "platform") %>%
  mutate(`Install Command` = sprintf(command_str, file)) %>%
  select(`Distribution`, `Download Link`, `Install Command`)

shinycannon_sh <- shinycannon_files %>%
  filter(platform == "sh") %>%
  unlist()

shinycannon_jar <- shinycannon_files %>%
  filter(platform == "jar") %>%
  unlist()
```

```{r echo=FALSE}
kable(linux_packages)
```

### macOS

1.  Install [Java](https://www.java.com/en/download/)
2.  Download `r shinycannon_sh[["Download Link"]]`
3.  Install with `sudo cp`r shinycannon_sh[["file"]]`/usr/local/bin/shinycannon; sudo chmod +x /usr/local/bin/shinycannon`
4.  Run `shinycannon -h` to see help output, or `shinycannon [RECORDING-PATH] [APP-URL]` to generate load.

### Windows

1.  Install [Java](https://www.java.com/en/download/)
2.  Download `r shinycannon_jar[["Download Link"]]` to the directory you wish to run it in
3.  Run `java -jar`r shinycannon_jar[["file"]]`-h` to see help output, or `java -jar`r shinycannon_jar[["file"]]`[RECORDING-PATH] [APP-URL]` to generate load.

## Recording

`shinycannon` accepts two required positional arguments:

1.  **[RECORDING-PATH]**: path to the file produced by `shinyloadtest::record_session`
2.  **[APP-URL]**: URL of the target Shiny application

In addition to these two required arguments, `shinycannon` accepts a number of optional arguments that can be specified with flags.
Of these, the most interesting are:

1.  **--workers**: The number of concurrent users to simulate. `shinycannon` uses threads to represent each user. It defaults to 1.
2.  **--loaded-duration-minutes**: The duration of the load test, in minutes. This does not include "warmup time", which is the time shinycannon spends gradually increasing the number of workers, or "cooldown time", which is the time spent decreasing the number of workers. It defaults to 0, meaning that after all workers have "warmed up", they will immediately begin to "cool down". Generally, you'll want to set this to a duration greater than 0. `shinycannon` will re-run the recording as necessary to fill the time.
3.  **--output-dir**: Name of the directory to create and store timing information in.

As an example, to run a load test simulating 5 concurrent users for at least 2 minutes, outputting to the directory `run1`:

``` {.bash}
$ shinycannon recording.log https://shinyapp.example.com/ --workers 5 --loaded-duration-minutes 2 --output-dir run1
2018-08-29 15:06:14.191 INFO [progress] - Running: 0, Failed: 0, Done: 0
2018-08-29 15:06:14.193 INFO [thread01] - Warming up
2018-08-29 15:06:14.195 INFO [thread00] - Waiting for warmup to complete
2018-08-29 15:06:19.193 INFO [progress] - Running: 1, Failed: 0, Done: 0
2018-08-29 15:06:24.194 INFO [progress] - Running: 1, Failed: 0, Done: 0
2018-08-29 15:06:29.083 INFO [thread02] - Warming up
2018-08-29 15:06:29.195 INFO [progress] - Running: 1, Failed: 0, Done: 0
2018-08-29 15:06:34.195 INFO [progress] - Running: 2, Failed: 0, Done: 0
2018-08-29 15:06:39.196 INFO [progress] - Running: 2, Failed: 0, Done: 0
2018-08-29 15:06:43.973 INFO [thread03] - Warming up
2018-08-29 15:06:44.196 INFO [progress] - Running: 2, Failed: 0, Done: 0
2018-08-29 15:06:49.196 INFO [progress] - Running: 3, Failed: 0, Done: 0
2018-08-29 15:06:54.201 INFO [progress] - Running: 3, Failed: 0, Done: 0
2018-08-29 15:06:58.862 INFO [thread04] - Warming up
2018-08-29 15:06:59.201 INFO [progress] - Running: 3, Failed: 0, Done: 0
2018-08-29 15:07:04.201 INFO [progress] - Running: 4, Failed: 0, Done: 0
2018-08-29 15:07:09.202 INFO [progress] - Running: 4, Failed: 0, Done: 0
2018-08-29 15:07:13.751 INFO [thread05] - Warming up
2018-08-29 15:07:13.752 INFO [thread00] - Maintaining for 2 minutes (120000 ms)
2018-08-29 15:07:14.202 INFO [progress] - Running: 4, Failed: 0, Done: 0
2018-08-29 15:07:19.202 INFO [progress] - Running: 5, Failed: 0, Done: 0
2018-08-29 15:07:24.202 INFO [progress] - Running: 5, Failed: 0, Done: 0
...
```

`shinycannon` includes detailed help documentation explaining the other arguments:

``` {.bash}
shinycannon -h
```

During the test, `shinycannon` reports the progress and number of simulated users.
The result of the test is an output directory (`run1` in the example above) which includes timing information for each session.
Inside that directory you'll find:

-   `recording.log`: a copy of the original recording used.
-   `detail.log`: log output, to help debug errors.
-   `./sessions/*.csv`: one file for each simulated user session.
