---
title: "Load Test Results"
output: 
  html_document: default
params: 
  url:
    label: "Deployed URL"
    value: "https://"
  concurrent:
    label: "Load Test: Number of Concurrent Connections"
    value: 4
  total: 
    label: "Load Test: Total Number of Connections"
    value: 4
  baseline:
    label: "Baseline Test: Total Number Sequential Connections"
    value: 4
  testFile:
    label: "Path to Load Test Script"
    value: "./tests/myloadtest.R"
  stagger: 
    label: "Maximum Delay in Staggering Concurrent Connections (sec)"
    value: 5
  loadTimeout:
    label: "Maximum time to wait for the Shiny app to load (sec)"
    value: 15
  phantomTimeout:
    label: "Maximum time to wait for phantomJS to start (sec)"
    value: 30
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(shinyloadtest)
library(ggplot2)
```

Load test run against `r params$url`. The load test target was **`r params$total` total visits** with **`r params$concurrent` concurrent users** at a time. Concurrent user sessions were staggered, with a max delay of `r params$stagger` seconds. Each "user" ran a series of actions against the application as recorded in `r params$testFile`.

```{r runTest, warning = FALSE}
loadTestLog <- loadTest(testFile = params$testFile, 
                        url = params$url,
                        numConcurrent = params$concurrent,
                        numTotal = params$total, 
                        loadTimeout = params$loadTimeout,
                        stagger = params$stagger,
                        phantomTimeout = params$phantomTimeout)
saveRDS(loadTestLog, file = "loadTestLog.Rds")
```

A baseline test was run with **`r params$baseline` total visits** occuring sequentially, one at a time.

```{r}
baselineLog <- loadTest(testFile = params$testFile, 
                        url = params$url,
                        numConcurrent = 1,
                        loadTimeout = params$loadTimeout,
                        numTotal = params$baseline, 
                        stagger = 0)
saveRDS(baselineLog, file = "baselineLog.Rds")
```


## Results

```{r}
errors <- getErrors(loadTestLog)
loadTestLog <- getSuccesses(loadTestLog)

baseline_errors <- getErrors(baselineLog)
baselineLog <- getSuccesses(baselineLog)

```


Of the targetted `r params$total` visits, `r length(unique(loadTestLog$connection))` were successful. 

```{r}
if (length(errors) > 0) {
  print("The following errors occured in the unsuccessful visits:")
  print(errors)
}

if (length(baseline_errors) > 0) {
  print("The following errors occured in the baseline test:")
  print(baseline_errors)
}

```

A maximum `r getMaxConcurrent(loadTestLog)` concurrent connections were achieved, compared to the target of `r params$concurrent`.

`r if ( length(errors) == 0 && (getMaxConcurrent(loadTestLog) < params$concurrent)) {message("Because all visits were successful, but the number of concurrent connections did not meet the target, consider lowering the stagger parameter or checking the server settings.")}`


## Comparison to Baseline

```{r}
load <- getPageLoadTimes(loadTestLog) 
load$type = "Under Load"

base <- getPageLoadTimes(baselineLog)
base$type = "Baseline"

results <- rbind(load, base)

ggplot(results, aes(load_time_sec, fill = type)) + 
  geom_density(alpha = 0.4) +
  theme_minimal() +
  labs(
    title = "Page Load Time",
    subtitle = paste(params$concurrent, "Concurrent Targetted"),
    x = "Load Time (sec)",
    y = "", 
    fill = ""
  )
```

The average page load time under load was `r mean(load$load_time_sec)` seconds compared to `r mean(base$load_time_sec)` seconds with no concurrent load.

The following graph shows the response time for specific events taken after page load. The table lists any events that may have timed out. Timeouts are expected for events that do not update an input. 

```{r}
load <- getSetInputTimes(loadTestLog)
base <- getSetInputTimes(baselineLog)

load$type = "Under Load"
base$type = "Basline"

results <- rbind(load, base)

## look for timeouts
timeout <- aggregate(results$timedout, by = list(id  = results$input_id), FUN = sum)
n <- length(unique(load$connection)) + length(unique(base$connection))

message <- paste0(timeout$id, ": ", timeout$x, " of ", n, " Timed Out")
data.frame(Warning = message, stringsAsFactors = FALSE)
```

**The distribution plots for events with timeouts should not be interpretted!**

```{r}
## plot the distributions
ggplot(results, aes(event_time_sec, fill =  type)) +
  geom_density(alpha = 0.4)  +
  facet_wrap(~input_id, scales = "free") +
  labs(
    title = "Application Usage",
    y = "",
    x = paste0("Event Time (sec)"),
    fill = ""
  ) +
  theme_minimal()
```

