---
title: "Load Test Results"
output: 
  html_document:
    df_print: paged
params: 
  directory:
    label: "Directory with profiles from load test"
    value: !r getwd()
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(shinyloadtest)
library(ggplot2)
```


```{r runTest, warning = FALSE}
log <- createLog(params$directory)
```


## Results

The load test was started at `r min(log$time)` and ended at `r max(log$time)`. There were `r length(unique(log$connection))` successful connections. A maximum `r getMaxConcurrent(log)` concurrent connections were achieved.

```{r concurrent_over_time}
con <- getConcurrentOverTest(log)
ggplot(con, aes(time, connections)) +
         geom_point() +
         geom_smooth() +
         theme_bw() +
         labs(
           title = "Concurrent connections over time"
         )

```

## Page load time

```{r}
load <- getPageLoadTimes(log) 

ggplot(load, aes(1, load_time_sec)) + 
  geom_boxplot() +
  theme_minimal() +
  scale_x_continuous(breaks = NULL) +
  labs(
    title = "Page Load Time",
    y = "Load Time (sec)",
    x = ""
  ) +
  guides(shape = FALSE)
```

The average page load time under load was `r mean(load$load_time_sec)` seconds.

The following graph shows the response time for specific events taken after page load.

```{r}
load <- getSetInputTimes(log)
load$inputId <- as.factor(load$inputId)
ggplot(load, aes(inputId, event_time_sec)) +
  geom_boxplot()  +
  labs(
    title = "Application Usage",
    x = "",
    y = paste0("Event Time (sec)"),
    fill = ""
  ) +
  theme_minimal() + 
  guides(fill = FALSE)
```

